Abstract

### 2025-05-23 output/Letter_to_editor_reviewers.Rmd
Address "dimensional sensitivity" comment

@@

>>>>

We acknowledge we may have overlooked this distinction.
To our understanding, a highly correlated latent space may imply a restricted dimensionality (as indicated by the rank of the covariance matrix) of the latent space.
We consider that this example, being an edge case (of unidimensionality), is very illustrative for our purposes, because it allows the reader to understand easily that all items should be aligned with the only dimension being measured and thus that grasp the intuitive notion that all direction cosines should be expected to be equal to 0.
However, in doing so, we have skipped a step in our reasoning, assimilating the dimensionality of the latent space with the dimensional sensitivity of the items.

To clarify better this point, we have tried to express better this distintion, emphasizing that the dimensionality of the latent space is an upper limit of the dimensional sensitivity of the items.
The aforementioned paragraph now reads (p. <!-- # TODO: Complete -->, lines <!-- # TODO: Complete -->):

> [...] In that example,
he uses a covariance matrix that practically has rank 1,
implying that the (alleged) three dimensions collapse into a single one.
This in turn makes the latent space unidimensional in practice.
As the dimensional sensitivity of the latent space sets the upper limit of
the dimensional sensitivity of a test measuring those traits,
it follows that the test must be unidimensional.
In such a case, when all the items are strictly aligned with
the one dimension they measure,
we should expect the direction angles to be extremely close to 0 for all items.
The reader interested can check that this is the case for
the covariance-based version of the direction angles
in the table in Online Resource 1.

@@

We are not completely sure to have understood this comment.
The discrimination values in the original `r MODEL_ACRONYM` metric are comparable, in the sense that they yield the same model probabilities for the same latent trait coordinates, independently of the covariance structure of the latent space.
However, the covariance-version of the multidimensional discrimination is different, and will indeed be dependent on the metric of the test space, which is defined by its geometrical basis, which determines the directions and scales of the axes.
In this case, the basis has been chosen to be a geometric representation of the correlation matrix (as given by the expression $`r INNER_PROD_CORR_COND`$, please see p. <!-- # TODO: Complete -->, line <!-- # TODO: Complete -->, in the new version), and therefore, the scales are not affected (only the relative directions among the axes).

We would kindly ask to elaborate further on this question if the Reviewer feels it has not been completely addressed.

### 2025-05-23 output/Main_text.Rmd
Address "dimensional sensitivity" comment

The latter effect can be better noticed in another case, in which Reckase
uses the same item parameters to demonstrate the over-specification of
dimensions (see p. 183).
In that example, he uses a covariance matrix that practically has rank 1,
implying that the (alleged) three dimensions collapse into a single one, being
thus the test unidimensional in practice.
In such a case, the covariance-based direction angles
are extremely close to 0 for all items, as one would expect
from a unidimensional test, where all items are strictly aligned with the one
dimension they measure. The reader interested can check out the parameters
computed in that case in the table in Online Resource 1.

>>>>

The latter effect can be better noticed in another case, in which Reckase
uses the same item parameters to demonstrate the over-specification of
dimensions (see p. 183).
In that example, he uses a covariance matrix that practically has rank 1,
implying that the (alleged) three dimensions collapse into a single one.
This in turn makes the latent space unidimensional in practice.
As the dimensional sensitivity of the latent space sets the upper limit of
the dimensional sensitivity of a test measuring those traits,
it follows that the test must be unidimensional.
In such a case, when all the items are strictly aligned with
the one dimension they measure,
we should expect the direction angles to be extremely close to 0 for all items.
The reader interested can check that this is the case for
the covariance-based version of the direction angles
in the table in Online Resource 1.
